### 2014-05-24 lineage-movement.tex
poorly worded meanderings about haplotypes within the patch

\paragraph{Questions:} 
\begin{enumerate}

>>>>

%%%%%%
\section{Haplotype lengths}

Some statements above about ``haplotype length'' aren't quite right.
Consider the haplotype shared about two B alleles;
by the time these coalesce, their ``shared haplotype'' in the above sense may be broken up into a number of blocks --
all coalescing linked to other B alleles.
More properly, the length of the contiguous haplotype
is determined by branching diffusions.
The single-lineage version is thus:
begin a diffusion with mass equal to genetic length (of something, rescaled);
then, any remaining particles split in two at rate proportional to their length,
always labeling the particle corresponding to the selected allele;
when this happens any particles other than the labeled particle may die
or may continue as an independent diffusion.

\section{Questions:} 
\begin{enumerate}

### 2014-05-24 patchy-selection-paper.tex
poorly worded meanderings about haplotypes within the patch

\begin{figure}[ht!]
  \begin{center}
    \includegraphics{sim-occupation-freqs}
  \end{center}
  \caption{
  \textbf{(A)} 
  Mean occupation frequencies of individual demes (grey dots)
  and averaged over demes at similar distances (black circles), on
  a one-dimensional grid of 201 demes, and
  a two-dimensional grid of $101\times 101$ demes.
  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq},
  with $C$ chosen arbitrarily to give a reasonable fit.
  (The uptick at the right end is due to boundary effects near the edge of the range.)
  \textbf{(B)} 
  Fluctuations: colored lines are snapshots of allele frequencies at evenly spaced times
  from a one-dimensional grid of 151 demes;
  the solid lines shows the cumulative mean occupation frequency across all generations.
  For both plots, the simulation was run for 1000 generations;
  the focal allele is beneficial in the central 20 demes, with $s_p=0.1$,
  and deleterious elsewhere, with $s_m=-0.02$; 
  demes had 1000 individuals in each;
  dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
 \gc{The approximation and simulations disagree at long distances from
   the patch, but this is due to edge effects in the simulations.} }   \label{fig:sim_occupation_freqs}
\end{figure}

@@

@@

First consider the lineage of a single B allele, moving back through time.
Suppose that the B allele is at mutation-selection balance, 
with the local proportion of B alleles near $x$ equal to $p(x)$,
and that we are close enough to the patch that fluctuations are small (unlike in previous sections).
If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
This means that in time $dt$, a total of $ dt \, \rho p(y) (1+s(y)) m(y,x) \, dy \, dx $ new, type B migrants 
will move from an area near $y$ of size $dy$ to an area near $x$ of size $dx$.
Looking backwards in time, this is the number of $B$ alleles living at $x$
who were produced by a parent from near $y$ in time $dt$.

XXX

>>>>

\begin{figure}[ht!]
  \begin{center}
    \includegraphics{sim-occupation-freqs}
  \end{center}
  \caption{
  \textbf{(A)} 
  Mean occupation frequencies of individual demes (grey dots)
  and averaged over demes at similar distances (black circles), on
  a one-dimensional grid of 201 demes, and
  a two-dimensional grid of $101\times 101$ demes.
  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq},
  with $C$ chosen arbitrarily to give a reasonable fit.
  (The uptick at the right end is due to boundary effects near the edge of the range.)
  \textbf{(B)} 
  Fluctuations: colored lines are snapshots of allele frequencies at evenly spaced times
  from a one-dimensional grid of 151 demes;
  the solid lines shows the cumulative mean occupation frequency across all generations.
  For both plots, the simulation was run for 1000 generations;
  the focal allele is beneficial in the central 20 demes, with $s_p=0.1$,
  and deleterious elsewhere, with $s_m=-0.02$; 
  demes had 1000 individuals in each;
  dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
  }   \label{fig:sim_occupation_freqs}
\end{figure}

@@

\plr{how's that?}

@@

First consider the lineage of a single B allele, moving back through time.
Suppose that the B allele is at mutation-selection balance, 
with the local proportion of B alleles near $x$ equal to $p(x)$,
and that we are close enough to the patch that fluctuations are small (unlike in previous sections).
If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
As argued by \citet{hudson},
in a discrete model the rate at which a B lineage moves from $x$ to $y$
is this influx divided by the local abundance of $B$ alleles:
\begin{align}
  r(x,y) = \frac{ p(y) (1+s(y)) m(y,x) }{ p(x) } .
\end{align}
After the same large-population, continuous-space limit that underlies the rest of this paper,
this motion should converge to a diffusion process \citep[that is similar to that studied by][]{hallatschek},
notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
In It\^o differential notation, 
the location $X_t$ of a B lineage $t$ generations ago satisfies
\begin{align}
  d X_t &= (1+s(X_t)) D \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,\\
    % &=  ( \log(p(X_t))' (1+s(X_t)) + s'(X_t) ) dt + \sqrt{1+s(X_t)} dB_t,\\
\end{align}
where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
$B_t$ is a standard Brownian motion, and $D$ denotes the gradient operator.
The main effect is the drift term, that pulls the lineage towards regions where $p(x)$ is larger;
since $s$ is generally small, the speed of the Brownian noise term has little effect.
Furthermore, it is easy to verify that the stationary distribution of this process is
\begin{align}
  \pi(x) = \frac{ p(x)^2 (1+s(x)) }{ \int_y p(y)^2 (1+s(y)) dy } .
\end{align}

If we are instead following a lineage at a locus at genetic distance $r$ from the selected locus,
and begin linked to a B allele,
then the lineage will move as a B lineage up until the most recent time
it is found linked to a b allele.
If the lineage is at $x$, the probability that a recombination that moves it from a B background
to a b background occurs in time $dt$ is $dt (1-p(x)) (1+s(x)/2)$.
The Feynman-Kac formula could then be used to write down differential equations
solved by various moments of the haplotype length as a function of the sampling location of the lineage.

More simply, consider: within the patch, where $p(x) \approx 1$, 
a linked lineage moves roughly as Brownian motion, and is unlikely to recombine onto a b background.
The rate of recombination between B and b backgrounds depends on the proportion of the patch in which $p(x)$ is intermediate.
This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
(and, therefore, the rate of recombination)
is roughly $r (\sigma/\sqrt{s})A^{-1/d}$.
This will be equal to the coalescence rate if
$1/(A\rho) = r (\sigma/\sqrt{s}) A^{-1/d}$, 
and hence the length of linkage shared within the patch is on the order of
\begin{align}
  \frac{1}{r} = \frac{ \sigma A^{(d-1)/d} }{\sqrt{s}} ,
\end{align}
which does not increase with $A$ in one dimension,
but increases with $\sqrt{A}$ in two dimensions.
